---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# inteRact

<!-- badges: start -->
<!-- badges: end -->

The goal of inteRact is to make affect control theory (ACT) equations accessible to a broader audience of social scientists. 

ACT is a theory of social behavior that hinges on the control principle of people acting in ways to confirm cultural meaning. ACT theoretical concepts have been used fruitfully in recent research within cultural sociology (Hunzaker 2014, 2018), stratification and occupational research (Freeland and Hoey 2018), social movements research (shuster and Campos-Castillo. 2017), gender and victimization (Boyle and McKinzie 2015; Boyle and Walker 2016). Information about ACT as a theory can be accessed here: https://research.franklin.uga.edu/act/ 

The goal of this package is to make elements of typical ACT analyses easier to implement into R: calculating the deflection of an event, the optimal behavior of an actor after an event, the relabeling of the actor or object after an event, and calculating emotions within events. You can look within the functions to see how the equations were programmed, but to truly get a handle on the equations and how they work, I refer you to Expressive Order (Heise 2010). 


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ekmaloney/inteRact")
```
## Example Analysis 

The following analysis is an example of how to use this package to implement ACT equations in your work. Functions can be applied to either a single event or a dataframe of events, to make analyzing a larger number of situations easier.

First, for any analysis you do, you need to ensure that the identities, behaviors, and modifiers that you study are measured in the ACT dictionary. inteRact includes the US dictionary that was measured in 2015 as the default dictionary (us_2015_full) and uses the US equations as default for calculations. 

Future versions will allow you to upload your own dictionaries and use them in calculations. 

The following code samples 20 identities and 10 behaviors to create a 10 event dataframe for the analysis. 

```{r example}
library(inteRact)
library(kableExtra)
library(knitr)
library(tibble)
library(tidyverse)

#load US dictionary 
data("us_2015_full")

#make a dataframe of events 
set.seed(814)
events <- tibble(actor = sample(us_2015_full$term[us_2015_full$type == "identity"], 10),
                 behavior = sample(us_2015_full$term[us_2015_full$type == "behavior"], 10),
                 object = sample(us_2015_full$term[us_2015_full$type == "identity"], 10))

kable(events, caption = "Sample Events for Analysis")
```

### Deflection 

To calculate deflection for a single event, you can use the function *calc_deflection*, which has 4 arguments: actor, behavior, object (all character strings), and dictionary (set to US currently). To calculate deflection for a series of events, you can use the function *batch_deflection* which has 1 argument: a dataframe with an actor, behavior, and object column (must be named that). For example, to find the deflection of `r print(paste(events$actor[1], " ", events$behavior[1], "s", " ", events$object[1], sep = ""))` you would do the following: 

```{r}
d <- get_deflection("brute", "work", "cook", equation = "us")

print(d)
```

To calculate the deflection of all of the events in the sample events dataframe, you would do the following: 
```{r}

events <- events %>% mutate(d = get_deflection(actor, behavior, object, equation = "us"))

kable(events)

```

### Element Deflection 

Additionally, you can see which elements contribute the most to the overall deflection by using the *element_deflection* function, as below. This returns a dataframe that includes each element's fundamental sentiment, the corresponding transient impression, the difference between the two, and the squared difference. In this way, you can see which element moved the most in EPA space after an event. In the case of brute works cook, the most deflection element-dimension is the Power of the Object (cook) followed by the Evaluation of the Actor (brute), 

```{r}
kable(element_deflection("brute", "work", "cook", equation = "us"), digits = 3)
```

You can do this for an entire dataframe with a little bit of nesting, as follows: 
```{r}

elem_def <- events %>%
            rowwise() %>%
            mutate(el_def = list(element_deflection(actor, behavior, object, equation = "us"))) %>%
            unnest(el_def)

kable(elem_def)
```
Following this, you can look at how different elements contribute to the overall deflection across your events: 
```{r}
ggplot(data = elem_def, mapping = aes(x = sqd_diff)) + 
      geom_histogram(binwidth = 0.5) + facet_grid(dimension ~ element) + 
      theme_minimal() + labs(x = "Squared Difference between Fundamental and Transient")
```

### Optimal Behavior 

ACT can also predict the optimal behavior for the actor to enact in the following interaction that would bring the transient sentiments back as close as possible to the fundamental sentiments. The function *optimal_behavior* finds the E, P, and A value for the optimal behavior

```{r}
beh <- optimal_behavior("brute", "work", "cook")
beh
```

### Closest Term 

To find the closest measured identity, modifier, or behavior to an EPA profile, you can use the function *closest_term* which has 5 arguments: the E, P, and A measurements, the type of term (identity, modifier, or behavior, and the maximum distance away in EPA units the term can be). 
```{r}
closest_term(beh$E, beh$P, beh$A, term_typ = "behavior", max_dist = 1)

closest_term(beh$E, beh$P, beh$A, term_typ = "behavior", max_dist = 0.5)
```
### Characteristic Emotion 

In ACT, there are three types of emotions: characteristic emotions, structural emotions, and consequent emotions. Characteristic emotions are the emotions that result from perfect confirmation of an identity - they are the emotions that allow the transient and fundamental impression to be equal. For example, the characteristic emotion of "brute" is -1.40, 3.05, 1.46 - moderately bad, very powerful, and moderately active. The closest modifier in EPA space is "domineering."

```{r}
ce <- characteristic_emotion("brute")
ce

closest_term(ce[1], ce[2], ce[3], term_typ = "modifier", max_dist = 1)
```

With a list of terms, you can get the EPA values for the characteristic emotions for all of them using the function *batch_characteristic_emotion*: 
```{r,  eval = FALSE, echo = FALSE}
batch_characteristic_emotion(events, var = "actor")
```

### Reidentification 
Following an ABO event, Actors and Objects can be reidentified by the other or an onlooker, to an identity that "fits" with their behavior better. The functions *reidentify_actor* and *reidentify_object* calculate this for you: 
```{r}

new_id <- events %>% 
          rowwise %>%  
          mutate(new_actor = list(reidentify_actor(actor, behavior, object, equation = "us"))) %>% 
          unnest(new_actor, names_sep = "_")

kable(new_id, digits = 2)

```

