---
title: "Example_Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(inteRact)
library(tidyverse)
library(here)
```

Step 1: Load data, select events

```{r}
#load US dict
data("us_2015_full")

#make a dataframe of events 
set.seed(729)
events <- tibble(actor = sample(us_2015_full$term[us_2015_full$type == "identity"], 50),
                 behavior = sample(us_2015_full$term[us_2015_full$type == "behavior"], 50),
                 object = sample(us_2015_full$term[us_2015_full$type == "identity"], 50))


head(events)
```
Step 2: New equation data 

-important -- equations should follow the "Z" format from Interact (Heise)


```{r}
your_equation_data <- read.table(here("data-raw/2010_equations.txt"))

your_equation_data

#format of your own eq data
colnames(your_equation_data)
your_equation_data$V1
```
Step 3: Reshape it to be in the necessary format 
- use the reshape_equation function on the "Z" formatted table to generate a dataframe with the necessary colnames and variables for the other functions in the package. 

```{r}
reshaped_eq <- reshape_new_equation(your_equation_data)


glimpse(reshaped_eq)
```

Step 4: run some calculations on the dataframe! 

If you have a df already set with your events -- you can format your code like this to calculate the deflection and get the optimal behavior for the 

```{r}

events %>% 
          rowwise() %>% 
          mutate(deflection = get_deflection(actor, behavior, object, 
                                             equation = "user supplied", 
                                             eq_df = reshaped_eq),
                 opt_beh = list(optimal_behavior(actor, behavior, object, 
                                                 equation = "user_supplied",
                                                 eq_df = reshaped_eq, which = "both"))) %>% 
          unnest(opt_beh) %>% head(10)

```

or if you only want predictions for actor/behavior --
```{r}

act_events <- events %>% 
          rowwise() %>% 
          mutate(deflection = get_deflection(actor, behavior, object, 
                                             equation = "user supplied", 
                                             eq_df = reshaped_eq),
                 opt_beh = list(optimal_behavior(actor, behavior, object, 
                                                 equation = "user_supplied",
                                                 eq_df = reshaped_eq, which = "actor"))) %>% 
  unnest(opt_beh)

act_events %>% 
          head(10)
```
You can also search for the closest behavior to the opt_behavior: 
```{r}
act_events <- act_events %>% 
              rowwise() %>% 
              mutate(closest_beh = closest_term(e = opt_E,
                                                p = opt_P,
                                                a = opt_A,
                                                term_typ = "behavior",
                                                max_dist = 1,
                                                num_terms = 1))

act_events %>% head()


```

